---
title: "Data Mining HW01"
output:
  pdf_document: default
  html_document: default
date: "2024-01-31"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE}
library(tidyverse)
library(lubridate)
library(ggplot2)
library(dplyr)
library(stringr)

# Load the dataset
ABIA <- read_csv("/Users/vita/Desktop/ABIA.csv")
  CRSDepTimegroup<-cut(ABIA$CRSDepTime, breaks = c(0, 100, 200, 300, 400, 500, 600, 700, 800, 900, 1000, 1100, 1200, 1300, 1400, 1500, 1600, 1700, 1800, 1900, 2000, 2100, 2200, 2300, Inf), 
     labels = c("0","1","2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12", "13", "14", "15", "16", "17", "18", "19", "20", "21", "22", "23"))

# Bind the groups to the main data frame
ABIA <- cbind(ABIA, CRSDepTimegroup)

# Summarize average departure delays
ABIA_sum = ABIA %>%
  drop_na(DepDelay) %>%
  group_by(CRSDepTimegroup) %>%
  summarise(avg_DepDelay = mean(DepDelay))

# Create a plot
ggplot(ABIA_sum, aes(x=CRSDepTimegroup, y=avg_DepDelay, group=1)) +
  geom_line(color="black") +
  geom_point(aes(color=avg_DepDelay), size=3) +
  scale_color_gradient(low="blue", high="red") +
  scale_x_discrete(labels = function(x) str_replace_all(x, "-", "\n")) +
  labs(x="Time of Day", y="Average Departure Delay (min)", title="Average Departure Delays by Time of Day") +
  theme_minimal()
```
```{r, echo=FALSE}
# What is the best time of day to fly to minimize delays, and does this change by airline?
# Best time is the early morning, before noon. It does change by airline. 

ABIA_sum <- ABIA %>%
  drop_na(DepDelay) %>%
  group_by(CRSDepTimegroup, UniqueCarrier) %>%
  summarise(avg_DepDelay = mean(DepDelay, na.rm = TRUE)) %>%
  ungroup() # ungroup for plotting

# Plot
ggplot(ABIA_sum, aes(x=CRSDepTimegroup, y=avg_DepDelay, group=UniqueCarrier, color=UniqueCarrier)) +
  geom_line() +
  geom_point(size=0.5) +
  scale_x_discrete(labels = function(x) str_replace_all(x, "-", "\n")) +
  labs(x="Time of Day", y="Average Departure Delay", title="Average Departure Delays by Time of Day and Airline") +
  theme_minimal() 
```
```{r, echo=FALSE}
ggplot(ABIA, aes(x = CRSDepTimegroup, y = DepDelay, fill = UniqueCarrier)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Average Departure Delays by Time of Day and Airline",
       x = "Time of Day",
       y = "Average Departure Delay (minutes)",
       fill = "Airline") +
  theme_minimal()
```
# What is the best time of year to fly to minimize delays, and does this change by destination?
# Yes, it changes by destination, and avoid summer & winter holidays.
```{r, echo=FALSE}
# Plot
ggplot(ABIA, aes(x = factor(Month), y = DepDelay, fill = Dest)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Average Departure Delays by Destination and Month",
       x = "Month",
       y = "Average Departure Delay (minutes)",
       fill = "Destination") +
  theme_minimal()
```
```{r, echo=FALSE}
library(tidyverse)
# Plot using ggplot2
ggplot(ABIA, aes(x = CRSDepTimegroup, y = DepDelay, fill = UniqueCarrier)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Average Departure Delays by Time of Day and Airline",
       x = "Time of Day",
       y = "Average Departure Delay (minutes)",
       fill = "Airline") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

# Q2
# A) What is the 95th percentile of heights for female competitors across all Athletics events (i.e., track and field)? Note that sport is the broad sport (e.g. Athletics) whereas event is the specific event (e.g. 100 meter sprint).
# The 95th percentile of heights for female competitors across all Athletics events is 183 cm.
```{r, echo=FALSE}
# Load the data
olympics <- read.csv("/Users/vita/Desktop/olympics_top20.csv")

# Calculate the 95th percentile of heights for female competitors in Athletics
percentile_95th_height <- olympics %>%
  filter(sex == 'F', sport == 'Athletics') %>%
  na.omit(height) %>% # Remove NA values for height
  summarise(percentile_95 = quantile(height, 0.95)) %>%
  .$percentile_95
print(percentile_95th_height)
```

# B) Which single women's event had the greatest variability in competitor's heights across the entire history of the Olympics, as measured by the standard deviation?
# The single women's event with the greatest variability in competitor's heights across the entire history of the Olympics, as measured by the standard deviation, is the Swimming Women's 100 metres Butterfly, with a standard deviation of approximately 8.13 cm.
```{r, echo=FALSE}
# Calculate standard deviation of heights for each event for female competitors
std_dev_heights <- olympics %>%
  filter(sex == 'F') %>%
  group_by(event) %>%
  summarise(std_dev_height = sd(height, na.rm = TRUE)) %>%
  arrange(desc(std_dev_height))

# Event with the greatest variability in heights
max_variability_event <- head(std_dev_heights, 20)
print(max_variability_event)
```

# C) How has the average age of Olympic swimmers changed over time? Does the trend look different for male swimmers relative to female swimmers?
# The average age: Both male and female swimmers average age has fluctuated over the years, with a slight increase in recent decades.
# Gender difference: There are periods where the average age of female swimmers appears to be slightly lower than that of male swimmers, but the gap between genders seems to have narrowed over time.

```{r, echo=FALSE}
# Calculate the average age of swimmers over time, separated by sex
avg_age_over_time <- olympics %>%
  filter(sport == 'Swimming') %>%
  group_by(year, sex) %>%
  summarise(average_age = mean(age, na.rm = TRUE))

# Plot
ggplot(avg_age_over_time, aes(x = year, y = average_age, color = sex)) +
  geom_line() +
  labs(title = "Average Age of Olympic Swimmers Over Time by Gender",
       x = "Year", y = "Average Age") +
  theme_minimal() 
```


